---
output: 
   github_document
editor_options: 
  chunk_output_type: console
---



<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
library(knitr)
library(rgl)
if (interactive()) setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
if (isTRUE(getOption('knitr.in.progress'))) options(rgl.useNULL=TRUE)
rgl::setupKnitr()
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  echo = TRUE, warning=FALSE, message=FALSE, include = TRUE, fig.show='hold'
)
library(gMOIP)
```

<!-- badges: start -->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/gMOIP)](https://CRAN.R-project.org/package=gMOIP)
[![CRAN_Downloads_Badge](http://cranlogs.r-pkg.org/badges/grand-total/gMOIP?color=brightgreen)](http://cranlogs.r-pkg.org/downloads/total/last-month/gMOIP)
[![R build status](https://github.com/relund/gMOIP/workflows/R-CMD-check/badge.svg)](https://github.com/relund/gMOIP/actions)
<!-- badges: end -->

# Tools for 2D and 3D plots of single and multi-objective linear/integer programming models

The `gMOIP` package can be used to make 2D and 3D plots of linear programming (LP), 
integer linear programming (ILP), or mixed integer linear programming (MILP) models 
with up to three objectives. Plots of both the solution and criterion space are possible.
For instance the non-dominated (Pareto) set for bi-objective LP/ILP/MILP programming models.

The package also include an `inHull` function for checking if a set of points is inside/at/outside the 
convex hull of a set of vertices (for arbitrary dimension).


## Usage

Consider the model $\max\{cx | Ax \leq b\}$ (could also be minimized) with 2 variables: 

```{r model}
A <- matrix(c(-3,2,2,4,9,10), ncol = 2, byrow = TRUE)
b <- c(3,27,90)
coeff <- c(7.75, 10)  # coefficients c
```

The polytope of the LP model with non-negative continuous variables ($x \geq 0$):

```{r lp}
plotPolytope(
   A,
   b,
   coeff,
   type = rep("c", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
```

The polytope of the ILP model with LP faces ($x\in \mathbb{Z}_0$):

```{r ilp}
plotPolytope(
   A,
   b,
   coeff,
   type = rep("i", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
```

The polytope of the MILP model (first variable integer) with LP faces:

```{r milp}
plotPolytope(
   A,
   b,
   coeff,
   type = c("i", "c"),
   crit = "max",
   faces = c("c", "c"),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
```

You can do the same with three variables:

```{r}
A <- matrix( c(
   3, 2, 5,
   2, 1, 1,
   1, 1, 3,
   5, 2, 4
), nc = 3, byrow = TRUE)
b <- c(55, 26, 30, 57)
coeff <- c(20, 10, 15)
```

```{r 3d, rgl=TRUE, include=TRUE}
# LP model
view <- matrix( c(-0.412063330411911, -0.228006735444069, 0.882166087627411, 0, 0.910147845745087,
                  -0.0574885793030262, 0.410274744033813, 0, -0.042830865830183, 0.97196090221405,
                  0.231208890676498, 0, 0, 0, 0, 1), nc = 4)   
loadView(v = view)  # set view angle
plotPolytope(A, b, plotOptimum = TRUE, obj = coeff, labels = "n")
```

The 3D plot is here a static image (so it can be seen in markdown). However, it may be an interactive plot as seen in [this example](http://htmlpreview.github.io/?https://github.com/relund/gMOIP/blob/master/inst/examples/3d_interactive.html).

For more examples see `example("gMOIP-package")` or `browseVignettes('gMOIP')`. 

## LaTeX support

You may create a TikZ file of the plot for LaTeX using

```{r, eval=FALSE}
library(tikzDevice)
tikz(file = "plot_polytope.tex", standAlone=F, width = 7, height = 6)
plotPolytope(
   A,
   b,
   coeff,
   type = rep("i", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
dev.off()
```

## Installation

Install the latest stable release from CRAN:

```{r, eval=FALSE}
install.packages("gMOIP")
```

Alternatively, install the latest development version from GitHub:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("relund/gMOIP")

library(gMOIP)
browseVignettes('gMOIP')
example("gMOIP-package")
```

