---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(gMOIP)
```


[![Travis-CI Build Status](https://travis-ci.org/relund/gMOIP.svg?branch=master)](https://travis-ci.org/relund/gMOIP)

# gMOIP - 2D plots of LP/IP programming models 

Make 2D plots of the polyeder of a linear programming (LP) or integer programming (IP) model,
including integer points and iso profit curve. Can also make a plot of a bi-objective criterion
space and the non-dominated (Pareto) set.

## Usage

```{r ex}
LP <- make.lp(0, 2)
set.objfn(LP, c(7.75, 10))
add.constraint(LP, c(9, 10), "<=", 90)
add.constraint(LP, c(2, 4), "<=", 27)
add.constraint(LP, c(-3, 2), "<=", 3)
colNames <- c("x1", "x2")
colnames(LP) <- colNames
set.type(LP, c(1,2), type = "integer")
control <- lp.control(LP, sense='max')

# Corner points of the polytope
cPoints<-cornerPoints(getA(LP), get.rhs(LP))
# Integer points in the polytope
iPoints<-integerPoints(LP)
# plot polytope (ggplot2)
plotPolytope(cPoints, iPoints, iso = getC(LP), crit = substr(lp.control(LP)$sense,1,3))
# Plot of criterion points given a bi-objective vector
zPoints<-criterionPoints(iPoints, c1 = c(getC(LP)[1], 0), c2 = c(0, getC(LP)[2]),
                         crit = substr(lp.control(LP)$sense,1,3))
plotCriterion(zPoints, addHull = FALSE, addTriangles = TRUE)
```

For more examples see `example("gMOIP-package")`.

## Installation

Install the latest stable release from CRAN:


```{r, eval=FALSE}
install.packages("gMOIP")
```

Alternatively, install the latest development version from GitHub:

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("relund/gMOIP")

library(gMOIP)
example("gMOIP-package")
```