---
title: "Plotting bi-objective models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting bi-objective models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(knitr)
library(rgl)
if (interactive()) setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
rgl::setupKnitr()
opts_chunk$set(
  collapse = TRUE,
  #cache = TRUE, autodep = TRUE, 
  comment = "#>",
  fig.show = "asis", 
  warning=FALSE, message=FALSE, include = TRUE, 
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)
```

With `gMOIP` you can also make plots of the criterion space for bi-objective models (linear
programming (LP), integer linear programming (ILP), or mixed integer linear programming (MILP)).
This vignette gives examples on how to make plots of both the solution and criterion space.

First we load the package:

```{r setup}
library(gMOIP)
```


## Bi-objective models with two variables

We define a function for grouping plots of the solution and criterion space:

```{r 2DPlotFunction}
plotBiObj2D <- function(A, b, obj,
   type = rep("c", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = FALSE,
   labels = "numb",
   addTriangles = TRUE,
   addHull = TRUE)
{
   p1 <- plotPolytope(A, b, type = type, crit = crit, faces = faces, plotFaces = plotFaces,
                      plotFeasible = plotFeasible, plotOptimum = plotOptimum, labels = labels) + 
      ggplot2::ggtitle("Solution space")
   p2 <- plotCriterion2D(A, b, obj, type = type, crit = crit, addTriangles = addTriangles,
                         addHull = addHull, plotFeasible = plotFeasible, labels = labels) +
      ggplot2::ggtitle("Criterion space")
   gridExtra::grid.arrange(p1, p2, nrow = 1) 
}
```

Let us define the constraints:

```{r 2DConstraints}
A <- matrix(c(-3,2,2,4,9,10), ncol = 2, byrow = TRUE)
b <- c(3,27,90)
```

First let us have a look at a LP model:

```{r 2DLP}
obj <- matrix(
   c(7, -10, # first criterion
     -10, -10), # second criterion
   nrow = 2)
plotBiObj2D(A, b, obj, addTriangles = FALSE)
```
