---
title: "Introduction to gMOIP"
author: "Lars Relund Nielsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to gMOIP}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(knitr)
library(rgl)
library(ggsci)
library(tidyverse)
library(magrittr)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
rgl::par3d("family" = "serif")
opts_chunk$set(
  collapse = TRUE,
  
  comment = "#>",
  warning=FALSE, message=FALSE, include = TRUE, 
  cache = TRUE,
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)
```

The `gMOIP` package can be used to make 2D and 3D plots of linear programming (LP), integer linear
programming (ILP), or mixed integer linear programming (MILP) models with up to three objectives.
This include the polytope, integer points, ranges and iso profit curve. Plots of both the solution
and criterion space are possible. For instance the non-dominated (Pareto) set for bi-objective
LP/ILP/MILP programming models.

The package also include an `inHull` function for checking if a set of points is inside/at/outside
the convex hull of a set of vertices (for arbitrary dimension).

Finally, the package also contains functions for generating (non-dominated) points in $\mathbb{R}_n$
and classifying non-dominated points as supported extreme, supported non-extreme and unsupported.


## Installation

Install the latest stable release from CRAN:

``` r
install.packages("gMOIP")
```

Alternatively, install the latest development version from GitHub:

``` r
install.packages("devtools")
devtools::install_github("relund/gMOIP")
library(gMOIP)
```

```{r, cache=FALSE, include=FALSE}
library(gMOIP)
```


## Single criterion models

We define the model \(\max\{cx | Ax \leq b\}\) (could also be minimized) using matrix `A` and
vectors `b` and `obj`:

```{r}
A <- matrix(c(-3,2,2,4,9,10), ncol = 2, byrow = TRUE)
b <- c(3,27,90)
obj <- c(7.75, 10)  # coefficients c
```

The polytope of the LP model with non-negative continuous variables
(\(x \geq 0\)):

```{r}
plotPolytope(
   A,
   b,
   obj,
   type = rep("c", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
```

The polytope of the ILP model with LP faces (\(x\in \mathbb{Z}_0\)):

```{r}
plotPolytope(
   A,
   b,
   obj,
   type = rep("i", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
```

The polytope of the MILP model (first variable integer) with LP faces:

```{r}
plotPolytope(
   A,
   b,
   obj,
   type = c("i", "c"),
   crit = "max",
   faces = c("c", "c"),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
```

### Three variables

You can do the same with three variables:

```{r}
A <- matrix( c(
   3, 2, 5,
   2, 1, 1,
   1, 1, 3,
   5, 2, 4
), nc = 3, byrow = TRUE)
b <- c(55, 26, 30, 57)
obj <- c(20, 10, 15)
```

LP model:

```{r, webgl=TRUE}
view <- matrix( c(-0.412063330411911, -0.228006735444069, 0.882166087627411, 0, 0.910147845745087,
                  -0.0574885793030262, 0.410274744033813, 0, -0.042830865830183, 0.97196090221405,
                  0.231208890676498, 0, 0, 0, 0, 1), nc = 4)   
loadView(v = view)  # set view angle
plotPolytope(A,
             b,
             obj,
             plotOptimum = TRUE,
             labels = "n")
```

ILP model (here with ILP faces):

```{r, webgl=TRUE}
view <- matrix( c(-0.412063330411911, -0.228006735444069, 0.882166087627411, 0, 0.910147845745087,
                  -0.0574885793030262, 0.410274744033813, 0, -0.042830865830183, 0.97196090221405,
                  0.231208890676498, 0, 0, 0, 0, 1), nc = 4)   
loadView(v = view)  # set view angle
plotPolytope(A,
             b,
             obj, 
             type = c("i", "i", "i"), 
             plotOptimum = TRUE, 
             labels = "n")
```

MILP model (here with continuous faces):

```{r, webgl=TRUE, cache=FALSE}
view <- matrix( c(-0.412063330411911, -0.228006735444069, 0.882166087627411, 0, 0.910147845745087,
                  -0.0574885793030262, 0.410274744033813, 0, -0.042830865830183, 0.97196090221405,
                  0.231208890676498, 0, 0, 0, 0, 1), nc = 4)   
loadView(v = view)  # set view angle
plotPolytope(A,
             b,
             obj, 
             type = c("i", "i", "c"), 
             faces = c("c", "c", "c"), 
             plotOptimum = TRUE, 
             # plotFaces = FALSE,
             labels = "n")
rglwidget()
```


## Bi-objective models

With `gMOIP` you can also make plots of the criterion space for bi-objective models (linear
programming (LP), integer linear programming (ILP), or mixed integer linear programming (MILP)).

First let us have a look at bi-objective model with two variables. We define a function for
grouping plots of the solution and criterion space:

```{r 2DPlotFunction}
plotBiObj2D <- function(A, b, obj,
   type = rep("c", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = FALSE,
   labels = "numb",
   addTriangles = TRUE,
   addHull = TRUE)
{
   p1 <- plotPolytope(A, b, type = type, crit = crit, faces = faces, plotFaces = plotFaces,
                      plotFeasible = plotFeasible, plotOptimum = plotOptimum, labels = labels) + 
      ggplot2::ggtitle("Solution space")
   p2 <- plotCriterion2D(A, b, obj, type = type, crit = crit, addTriangles = addTriangles,
                         addHull = addHull, plotFeasible = plotFeasible, labels = labels) +
      ggplot2::ggtitle("Criterion space")
   gridExtra::grid.arrange(p1, p2, nrow = 1) 
}
```

Let us define the constraints:

```{r 2DConstraints}
A <- matrix(c(-3,2,2,4,9,10), ncol = 2, byrow = TRUE)
b <- c(3,27,90)
```

First let us have a look at a LP model (maximize):

```{r 2DLP}
obj <- matrix(
   c(7, -10, # first criterion
     -10, -10), # second criterion
   nrow = 2)
plotBiObj2D(A, b, obj, addTriangles = FALSE)
```

Note the non-dominated (Pareto) set consists of all supported extreme non-dominated points
(illustrated with triangles) and the line segments between them.

ILP model (maximize):

```{r 2DILPMax}
obj <- matrix(c(3, -1, -2, 2), nrow = 2)
plotBiObj2D(A, b, obj, type = rep("i", ncol(A)))
```

Note the non-dominated set consists of all points in black (with shape supported extreme:triangle,
supported non-extreme:round, unsupported:round (not on the border)). The triangles drawn using the
extreme non-dominated points illustrate areas where unsupported non-dominated points may be found.
A point in the solution space is identified in the criterion space using the same number.

ILP models (minimize):

```{r 2DILPMin}
obj <- matrix(c(3, -1, -2, 2), nrow = 2)
plotBiObj2D(A, b, obj, type = rep("i", ncol(A)), crit = "min")
```


MILP model ($x_2$ integer) (minimize):

```{r 2DMILPMin}
obj <- matrix(c(3, -1, -2, 2), nrow = 2)
plotBiObj2D(A, b, obj, type = c("c", "i"), crit = "min")
```

Note the solution space now consists to segments and hence the non-dominated set may consist of
points and segments (open and closed). Note these segments are not highlighted in the current 
version of `gMOIP`.


## Bi-objective models with three variables

We define functions for plotting the solution and criterion space:

```{r 3DPlotFunction}
plotSol <- function(A, b, type = rep("c", ncol(A)),
                        faces = rep("c", ncol(A)),
                        plotFaces = TRUE, labels = "numb")
{
   loadView(v = view, close = F, zoom = 0.75)
   plotPolytope(A, b, type = type, faces = faces, labels = labels, plotFaces = plotFaces, 
                argsTitle3d = list(main = "Solution space"))
}

plotCrit <- function(A, b, obj, crit = "min", type = rep("c", ncol(A)), addTriangles = TRUE, 
                     labels = "numb") 
{
    plotCriterion2D(A, b, obj, type = type, crit = crit, addTriangles = addTriangles, 
                   labels = labels) + 
      ggplot2::ggtitle("Criterion space")
}
```

Four examples are given. A few plots of the solution space are made interactive to illustrate the
functionality.

### Example 1

We define the model  $\max \{cx | Ax \leq b\}$ (could also be minimized) with three variables:

```{r Ex1Model}
Ab <- matrix( c(
   1, 1, 2, 5,
   2, -1, 0, 3,
   -1, 2, 1, 3,
   0, -3, 5, 2
), nc = 4, byrow = TRUE)
A <- Ab[,1:3]
b <- Ab[,4]
obj <- matrix(c(1, -6, 3, -4, 1, 6), nrow = 2)
```

We load the preferred view angle for the RGL window:

```{r Ex1View}
view <- matrix( c(-0.452365815639496, -0.446501553058624, 0.77201122045517, 0, 0.886364221572876,
                  -0.320795893669128, 0.333835482597351, 0, 0.0986008867621422, 0.835299551486969,
                  0.540881276130676, 0, 0, 0, 0, 1), nc = 4)
loadView(v = view)
```

LP model (solution space):

```{r Ex1LPSol, webgl=TRUE, cache=FALSE}
plotSol(A, b)
```

LP model (criterion space):

```{r Ex1LPCrit}
plotCrit(A, b, obj, addTriangles = FALSE) 
```

ILP model (solution space):

```{r Ex1ILPSol, webgl=TRUE, cache=FALSE}
plotSol(A, b, type = c("i","i","i"))
```

ILP model (criterion space):

```{r Ex1ILPCrit}
plotCrit(A, b, obj, type = c("i","i","i"))
```

### Example 2

```{r Ex3Model}
A <- matrix( c(
   1, 1, 1,
   3, 0, 1
), nc = 3, byrow = TRUE)
b <- c(10, 24)
obj <- matrix(c(1, -6, 3, -4, 1, 6), nrow = 2)
```

We load the preferred view angle for the RGL window:

```{r Ex3View}
view <- matrix( c(-0.812462985515594, -0.029454167932272, 0.582268416881561, 0, 0.579295456409454,
                  -0.153386667370796, 0.800555109977722, 0, 0.0657325685024261, 0.987727105617523,
                  0.14168381690979, 0, 0, 0, 0, 1), nc = 4)
loadView(v = view)
```

LP model (solution space):

```{r Ex3LPSol, webgl=TRUE, cache=FALSE}
plotSol(A, b)
```

LP model (criterion space):

```{r Ex3LPCrit}
plotCrit(A, b, obj, addTriangles = FALSE) 
```


MILP model with variable 2 integer (solution space):

```{r Ex3MILPSol5, webgl=TRUE, cache=FALSE}
plotSol(A, b, type = c("c","i","c"), plotFaces = F)
```

MILP model with variable 2 integer (criterion space):

```{r Ex3MILPCrit5}
plotCrit(A, b, obj, type = c("c","i","c"))
```


## Tri-objective models

The criterion space can be plotted for tri-objective models. 

Example with many unsupported:

```{r, webgl=TRUE, cache=FALSE}
view <- matrix( c(0.333316594362259, 0.938472270965576, -0.0903875231742859, 0, 0.83994072675705, -0.339126199483871, -0.423665106296539, 0, -0.428250730037689, 0.0652943551540375, -0.901297807693481, 0, 0, 0, 0, 1), nc = 4)
loadView(v = view)
pts <- genNDSet(3, 50, argsSphere = list(below = FALSE), dubND = FALSE)
pts <- classifyNDSet(pts[,1:3])
head(pts)
ini3D(argsPlot3d = list(xlim = c(min(pts[,1])-2,max(pts[,1])+2),
   ylim = c(min(pts[,2])-2,max(pts[,2])+2),
   zlim = c(min(pts[,3])-2,max(pts[,3])+2)))
plotPoints3D(pts[pts$se,1:3], argsPlot3d = list(col = "red"))
plotPoints3D(pts[!pts$sne,1:3], argsPlot3d = list(col = "black"))
plotPoints3D(pts[!pts$us,1:3], argsPlot3d = list(col = "blue"))
plotCones3D(pts[,1:3], rectangle = TRUE, argsPolygon3d = list(alpha = 1, color = "cornflowerblue"))
plotHull3D(pts[,1:3], addRays = TRUE, argsPolygon3d = list(alpha = 0.25, color = "red"), useRGLBBox = TRUE)
finalize3D(argsAxes3d = list(edges = "bbox"))
```

Example with many supported:

```{r, webgl=TRUE, cache=FALSE}
loadView(v = view)
pts <- genNDSet(3, 50, argsSphere = list(below = TRUE), dubND = FALSE)
pts <- classifyNDSet(pts[,1:3])
ini3D(argsPlot3d = list(xlim = c(min(pts[,1])-2,max(pts[,1])+2),
   ylim = c(min(pts[,2])-2,max(pts[,2])+2),
   zlim = c(min(pts[,3])-2,max(pts[,3])+2)))
plotPoints3D(pts[pts$se,1:3], argsPlot3d = list(col = "red"))
plotPoints3D(pts[!pts$sne,1:3], argsPlot3d = list(col = "black"))
plotPoints3D(pts[!pts$us,1:3], argsPlot3d = list(col = "blue"))
plotCones3D(pts[,1:3], rectangle = TRUE, argsPolygon3d = list(alpha = 1, color = "cornflowerblue"))
plotHull3D(pts[,1:3], addRays = TRUE, argsPolygon3d = list(alpha = 0.25, color = "red"), useRGLBBox = TRUE)
finalize3D(argsAxes3d = list(edges = "bbox"))
```


## Drawing shapes related to MO

```{r color scales, include=FALSE}
mypal <- pal_npg("nrc", alpha = 0.7)(9)
mypal <- pal_simpsons()(9)
# mypal
library("scales")
# show_col(mypal)
```

Complex shapes can be drawn using cones, hulls etc. Plot of a single non-dominated point and its corresponding local upper bounds given the bounding box. Green areas are overlapping if branch on each local upper bound.

```{r Fig3 in Forget20, webgl = TRUE, echo=FALSE, cache=FALSE}
ubSet <- matrix( c(1,1,1), ncol = 3, byrow = TRUE)
row.names(ubSet) <- paste0("z", 1)
lim <- c(0,7)  # bounding box for problem
m <- rep(lim[1],3)
M <-rep(lim[2],3)
augment <- function(idx, a, b, prefix = "") {
  res <- NULL
  for (i in idx) {
    c <- b
    c[i] <- a[i]
    res <- rbind(res,c)
  }
  row.names(res) <- paste0(prefix,idx)
  return(res)
}
zHat <- augment(1:3, M, m)
# iteration 0
U <- matrix(M, ncol = 3, byrow = T)
row.names(U) <- "u0"
N <- NULL
NHat <- rbind(N,zHat)
z <- ubSet[1,1:3]
u0 <- M
U <- augment(1:3, z, u0, "u")
N<- rbind(N, ubSet[1,1:3, drop=F])

view <- matrix( c(-0.640044808387756, -0.205455526709557, 0.740358412265778, 0, 0.767741799354553, -0.208960145711899, 0.605729579925537, 0, 0.0302549730986357, 0.956098079681396, 0.29148057103157, 0, 0, 0, 0, 1), nc = 4)

rgl::mfrow3d(nr = 1, nc = 2, sharedMouse = F)

lim <- c(lim[1]-0.05,lim[2]+0.05)
## First plot
ini3D(argsPlot3d = list(xlim = lim, ylim = lim, zlim = lim))
# plotRectangle3D(m,M)
plotPoints3D(N, addText = "coord", argsPlot3d = list(type="p", col = "black", size = 8))
plotPoints3D(U, addText = "coord", argsPlot3d = list(type="p", col = mypal[3], size = 8), argsText3d = list(col = "blue", adj = c(1,1)))
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPersp3d = list(col = mypal[2]), usePersp3d = T, lines = 50)
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPlanes3d = list(alpha = 0.7, col = mypal[8]))
plotCones3D(U[1:3,], drawPoint = F, argsPolygon3d = list(alpha =0.3, color = mypal[1]), rectangle = F, direction = -1, drawLines = F)
finalize3D(argsAxes3d = list(edges = "bbox", box = F))
loadView(v = view, clear = FALSE)

## all relevant points for drawing search areas
vtx <- matrix(c(
  1,1,4,
  1,4,1,
  4,1,1,
  0,0,7,
  0,7,0,
  7,0,0,
  0,0,6,
  0,6,0,
  6,0,0,
  0,1,5,
  0,5,1,
  1,0,5,
  1,5,0,
  5,0,1,
  5,1,0,
  1,1,7,
  1,7,1,
  7,1,1,
  7,1,0,
  7,0,1,
  1,7,0,
  1,0,7,
  0,7,1,
  0,1,7,
  7,7,0,
  7,0,7,
  0,7,7,
  7,1,7,
  7,7,1,
  1,7,7
), ncol = 3, byrow = T)
vtx <- vtx %>% set_colnames(c("x", "y", "z")) %>% as_tibble() %>% arrange(x,y,z)
# plotPoints3D(vtx, addText = "coord")
# print(vtx, n=100)

## Second plot
ini3D(clear = FALSE, argsPlot3d = list(xlim = lim, ylim = lim, zlim = lim))
plotPoints3D(N, addText = "coord", argsPlot3d = list(type="p", col = "black", size = 8))
plotPoints3D(U, addText = "coord", argsPlot3d = list(type="p", col = mypal[7], size = 8), argsText3d = list(col = "blue", adj = c(1,1)))
# plotPlane3D(c(1,1,1), point = c(2,2,2), argsPersp3d = list(col = mypal[2]), usePersp3d = T, lines = 50)
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPlanes3d = list(alpha = 0.3, col = mypal[8]))

args <- list(color = mypal[9], alpha = 0.7)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,x,z)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(z,y,x)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)

args <- list(color = mypal[4], alpha = 0.2)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,x,z)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,z,x)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
finalize3D(argsAxes3d = list(edges = "bbox", box = F))
loadView(v = view, clear = FALSE)
```

A set of non-dominated points and the areas they dominate. Light blue plane are a lower bound.

```{r Fig4, webgl = TRUE, echo=FALSE, cache=FALSE}
pts <- data.frame(
          z1 = c(-2, -2, -2, -1, -1, 0, 0, 1, 1, 2, 3, 4, 4, 5, 5, 6, 6, 6),
          z2 = c(2, 4, 5, 1, 6, 0, 6, -1, 6, 5, -2, -2, 4, -2, 3, -1, 0, 2),
          z3 = c(5, 4, 3, 6, 2, 6, 0, 6, -1, -2, 5, 4, -2, 3, -2, 1, 0, -1),
          nd = c(TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,
                 TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE)
)

## First plot
open3d()
mfrow3d(nr = 1, nc = 2, sharedMouse = F)
ini3D(argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10),
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10),
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
lub <- Rfast::colMinsMaxs(as.matrix(pts[,1:3]))[2,]
M <- rgl::par3d()$bbox
lub <- rbind(lub, c(M[2],M[4],M[6])-0.01)
llb <- c(M[1],M[3],M[5])+ 0.01
# plotPoints3D(pts)
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T, rectangle = T)
plotCones3D(lub[2,], direction = -1, argsPolygon3d = list(alpha = 0.4, color = mypal[7]), drawPoint = F)
# plotHull3D(unique(pts))
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.4, col = mypal[7]))
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(llb, argsPolygon3d = list(alpha = 0.4, color = mypal[4]), drawPoint = F)
# plotHull3D(unique(pts))
clipplanes3d(1, 1, 1, -9)
useSubscene3d(root)
axes3d()

## Second plot
ini3D(clear = FALSE, argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10),
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10),
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
# plotPoints3D(pts)
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.8, col = mypal[7]))
plotCones3D(lub[1,], direction = -1, argsPolygon3d = list(alpha = 1, color = mypal[4]), drawPoint = T)
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T, drawLines = F)
useSubscene3d(root)
axes3d()
```


## Generating points

Points in $\mathbb{R}_n$ can be generated using function `genSample`. Different methods can be used for generation:

**Random**

The coefficient are generated randomly with a uniform distribution in the range $[a,b]$. For three objectives, random generated coefficients looks as follows.

```{r, webgl = TRUE, cache=FALSE}
range <- matrix(c(1,100, 50,100, 10,50), ncol = 2, byrow = TRUE )
open3d()
mfrow3d(nr = 1, nc = 1)
pts <- genSample(3, 1000, range = range, random = TRUE)
plotPoints3D(pts)
finalize3D()
```

**Sphere down**

The coefficients are generated on the lower part of a sphere (see next picture). Note that the sphere is adjusted such that the coefficients are in the range $[a,b]$, i.e. the sphere is not necessarily included in $[a,b]^p$.

```{r, webgl = TRUE, cache=FALSE}
cent <- c(1000,1000,1000)
r <- 750
planeC <- c(cent-r/3)
planeC <- c(planeC, -sum(planeC^2))
pts <- genSample(3, 500,
  argsSphere = list(center = cent, radius = r, below = TRUE, plane = planeC, factor = 6))
ini3D()
rgl::spheres3d(cent, radius=r, color = "grey100", alpha=0.1)
plotPoints3D(pts)
rgl::planes3d(planeC[1],planeC[2],planeC[3],planeC[4], alpha = 0.5, col = "red")
finalize3D()
```

**Sphere up**

The coefficients are generated on the upper part of a sphere (see next picture). Note that the sphere is adjusted such that the coefficients are in the range $[a,b]$, i.e. the sphere is not necessarily included in $[a,b]^p$.

```{r, webgl = TRUE, cache=FALSE}
cent <- c(1000,1000,1000)
r <- 750
planeC <- c(cent+r/3)
planeC <- c(planeC, -sum(planeC^2))
pts <- genSample(3, 500,
  argsSphere = list(center = cent, radius = r, below = FALSE, plane = planeC, factor = 6))
ini3D()
rgl::spheres3d(cent, radius=r, color = "grey100", alpha=0.1)
plotPoints3D(pts)
rgl::planes3d(planeC[1],planeC[2],planeC[3],planeC[4], alpha = 0.5, col = "red")
finalize3D()
```

**2box**

The coefficients are generated randomly but in two specific parts of $[a,b]^p$ (see next picture).

```{r, webgl = TRUE, cache=FALSE}
range <- matrix(c(1,1000, 1,1000, 1,1000), ncol = 2, byrow = TRUE )
ini3D(argsPlot3d = list(box = TRUE, axes = TRUE))
pts <- genSample(3, 300, range = range, box = TRUE)
plotPoints3D(pts)
finalize3D()
```

## Classifying

Non-dominated points can be classified using `classifyNDSet`:

```{r}
pts <- matrix(c(0,0,1, 0,1,0, 1,0,0, 0.5,0.2,0.5, 0.25,0.5,0.25), ncol = 3, byrow = TRUE)
open3d()
ini3D(argsPlot3d = list(xlim = c(min(pts[,1])-2,max(pts[,1])+2),
  ylim = c(min(pts[,2])-2,max(pts[,2])+2),
  zlim = c(min(pts[,3])-2,max(pts[,3])+2)))
plotHull3D(pts, addRays = TRUE, argsPolygon3d = list(alpha = 0.5), useRGLBBox = TRUE)
pts <- classifyNDSet(pts[,1:3])
plotPoints3D(pts[pts$se,1:3], argsPlot3d = list(col = "red"))
plotPoints3D(pts[!pts$sne,1:3], argsPlot3d = list(col = "black"))
plotPoints3D(pts[!pts$us,1:3], argsPlot3d = list(col = "blue"))
plotCones3D(pts[,1:3], rectangle = TRUE, argsPolygon3d = list(alpha = 1))
finalize3D()
pts

pts <- genNDSet(3,50, dubND = FALSE)[,1:3]
open3d()
ini3D(argsPlot3d = list(xlim = c(0,max(pts$z1)+2),
  ylim = c(0,max(pts$z2)+2),
  zlim = c(0,max(pts$z3)+2)))
plotHull3D(pts, addRays = TRUE, argsPolygon3d = list(alpha = 0.5))
pts <- classifyNDSet(pts[,1:3])
plotPoints3D(pts[pts$se,1:3], argsPlot3d = list(col = "red"))
plotPoints3D(pts[!pts$sne,1:3], argsPlot3d = list(col = "black"))
plotPoints3D(pts[!pts$us,1:3], argsPlot3d = list(col = "blue"))
finalize3D()
pts
```

The classification is done using the qhull algorithm that find the convex hull of the points including the rays. If a vertex then if must be supported extreme. Next we use the `inHull` algorithm to find out if the remaining are on the border or not (supported non-extreme and unsupported). 

## Animations

Animations may also be done (see example).


## LaTeX support

You may create a TikZ file of the 2D plots for LaTeX using

``` r
library(tikzDevice)
tikz(file = "plot_polytope.tex", standAlone=F, width = 7, height = 6)
plotPolytope(
   A,
   b,
   obj,
   type = rep("i", ncol(A)),
   crit = "max",
   faces = rep("c", ncol(A)),
   plotFaces = TRUE,
   plotFeasible = TRUE,
   plotOptimum = TRUE,
   labels = "coord"
)
dev.off()
```



## Further examples

For further examples see the documentation and the vignettes.

``` r
library(gMOIP)
browseVignettes('gMOIP')
example("gMOIP-package")
```



