---
title: "Generating costs in multi-objective programming"
author: "Lars Relund Nielsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generating costs in multi-objective programming}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
library(knitr)
library(rgl)
library(ggsci)
library(magrittr)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
#rgl::par3d("family" = "serif")
opts_chunk$set(
  collapse = TRUE,
  warning=FALSE, message=FALSE, include = TRUE, 
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)

if (!requireNamespace("rmarkdown", quietly = TRUE) || !rmarkdown::pandoc_available("1.14")) {
   warning(call. = FALSE, "These vignettes assume rmarkdown and pandoc version 1.14 (or higher). These were not found. Older versions will not work.")
   knitr::knit_exit()
}
```

When generating instances for testing in multi-objective programming the cost structure of the objective coefficients may have a huge impact on the difficulty of the instance. Points in $\mathbb{R}_n$ can be generated using function `genSample`. Different methods can be used for generation:

**Random**

The coefficient are generated randomly with a uniform distribution in the range $[a,b]$. For three objectives, random generated coefficients looks as follows.

```{r}
library(gMOIP)
range <- matrix(c(1,100, 50,100, 10,50), ncol = 2, byrow = TRUE )
pts <- genSample(3, 1000, range = range, random = TRUE)
ini3D()
plotPoints3D(pts)
finalize3D()
rglwidget()
```

**Sphere down**

The coefficients are generated on the lower part of a sphere (see next picture). Note that the sphere is adjusted such that the coefficients are in the range $[a,b]$, i.e. the sphere is not necessarily included in $[a,b]^p$.

```{r, eval=T}
cent <- c(1000,1000,1000)
r <- 750
planeC <- c(cent-r/3)
planeC <- c(planeC, -sum(planeC^2))
pts <- genSample(3, 500,
  argsSphere = list(center = cent, radius = r, below = TRUE, plane = planeC, factor = 6))
ini3D()
plotPoints3D(pts)
planes3d(planeC[1],planeC[2],planeC[3],planeC[4], alpha = 0.5, col = "red")
spheres3d(cent, radius=r, color = "grey100", alpha=0.1)
finalize3D()
rglwidget(reuse = F)
```

**Sphere up**

The coefficients are generated on the upper part of a sphere (see next picture). Note that the sphere is adjusted such that the coefficients are in the range $[a,b]$, i.e. the sphere is not necessarily included in $[a,b]^p$.

```{r}
cent <- c(1000,1000,1000)
r <- 750
planeC <- c(cent+r/3)
planeC <- c(planeC, -sum(planeC^2))
pts <- genSample(3, 500,
  argsSphere = list(center = cent, radius = r, below = FALSE, plane = planeC, factor = 6))
ini3D()
spheres3d(cent, radius=r, color = "grey100", alpha=0.1)
plotPoints3D(pts)
planes3d(planeC[1],planeC[2],planeC[3],planeC[4], alpha = 0.5, col = "red")
finalize3D()
rglwidget(reuse = F)
```

**2box**

The coefficients are generated randomly but in two specific parts of $[a,b]^p$ (see next picture).

```{r, webgl = TRUE, cache=FALSE}
range <- matrix(c(1,1000, 1,1000, 1,1000), ncol = 2, byrow = TRUE )
ini3D(argsPlot3d = list(box = TRUE, axes = TRUE))
pts <- genSample(3, 300, range = range, box = TRUE)
plotPoints3D(pts)
finalize3D()
```

```{r, include=F}
rm(list = ls(all.names = TRUE))
```