---
title: "Shapes in multi-objective programming"
author: "Lars Relund Nielsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shapes in multi-objective programming}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(knitr)
library(rgl)
library(ggsci)
library(ggplot2)
library(tibble)
library(dplyr)
library(magrittr)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
rgl::par3d("family" = "serif")
opts_chunk$set(
  collapse = TRUE,
  echo = FALSE, 
  comment = "#>",
  warning=FALSE, message=FALSE, include = TRUE, 
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)
set.seed(78974)

if (!requireNamespace("rmarkdown", quietly = TRUE) || !rmarkdown::pandoc_available("1.14")) {
   warning(call. = FALSE, "These vignettes assume rmarkdown and pandoc version 1.14 (or higher). These were not found. Older versions will not work.")
   knitr::knit_exit()
}
```

Package `gMOIP` has different features for plotting different shapes in 3D. 

```{r}
library(gMOIP)
```

We define the sets $\mathbb{R}_{>} = \{ z\in \mathbb{R} \mid z > 0\}$, $\mathbb{R}_{\geqq} = \{ z\in \mathbb{R} \mid z \geqq 0\}$ and $\mathbb{R}_{\geqslant} = \{ z\in \mathbb{R} \mid z \geqslant 0\}$:

```{r}
rgl::mfrow3d(nr = 1, nc = 3, sharedMouse = TRUE)
## Set R^3_>=
ini3D(argsPlot3d = list(xlim = c(-5, 100), ylim = c(-5, 100), zlim = c(-5, 100)))
plotCones3D(c(0,0,0), drawPoint = F, argsPolygon3d = list(alpha=0.8))
title3d(main = "R_>=")
finalize3D()
next3d()
## Set R^3_>-
ini3D(clear = FALSE, argsPlot3d = list(xlim = c(-5, 100), ylim = c(-5, 100), zlim = c(-5, 100)))
plotCones3D(c(0,0,0), drawPoint = F, argsPolygon3d = list(alpha=0.8))
rgl::spheres3d(0,0,0, radius = 2, alpha = 0.8, col="gray80")
title3d(main = "R_>-")
finalize3D()
next3d()
## Set R^3_>
ini3D(clear = FALSE, argsPlot3d = list(xlim = c(-5, 100), ylim = c(-5, 100), zlim = c(-5, 100)))
plotCones3D(c(0,0,0), drawPoint = F, argsPolygon3d = list(alpha=0.3, color = "red"))
title3d(main = "R_>")
finalize3D()
rglwidget()
close3d()
```

The non-dominated set (red points are non-dominated and black dominated): 

```{r nd points}
library(gMOIP)
pts <- genNDSet(3, 100, keep = TRUE, dubND = FALSE)
ini3D()
plotPoints3D(pts[!pts$nd,1:3])
plotPoints3D(pts[pts$nd,1:3], argsPlot3d = list(col = "red"))
finalize3D()
rglwidget()
```


We may also add dominance cones and the hull:

```{r addDCones}
lim <- as.data.frame(apply(pts, 2, function(x) c(min(x)-5, max(x)+5)))
ini3D(argsPlot3d = list(xlim = lim$z1, ylim = lim$z2, zlim = lim$z3))
plotHull3D(
  pts[pts$nd, 1:3],
  addRays = T,
  drawPoints = T,
  argsPlot3d = list(col = "red"),
  argsPolygon3d = list(color = "red", alpha = 0.5)
)
plotCones3D(pts[pts$nd,1:3], argsPolygon3d = list(color = "grey100", alpha = 1), rectangle = TRUE)
finalize3D()
rglwidget()
```

Note the area between the hull and the cones is the search area where further nondominated points may be found.

Another example with more unsupported points:

```{r}
pts1 <- genNDSet(3, 100, keep = T, argsSphere = list(below = F))
lim1 <- as.data.frame(apply(pts1, 2, function(x) c(min(x)-5, max(x)+5)))
ini3D(argsPlot3d = list(xlim = lim1$z1, ylim = lim1$z2, zlim = lim1$z3))
plotPoints3D(pts1[!pts1$nd, 1:3])
plotHull3D(
  pts1[pts1$nd, 1:3],
  addRays = T,
  drawPoints = T,
  argsPlot3d = list(col = "red"),
  argsPolygon3d = list(color = "red", alpha = 0.5)
)
plotCones3D(pts1[pts1$nd, 1:3],
            argsPolygon3d = list(color = "grey100", alpha = 1),
            rectangle = T)
finalize3D()
rglwidget()
```

We may also plot reverse dominance cones:

```{r}
pts2 <- genNDSet(3, 50, range = c(1,20), dubND = FALSE)
lim2 <- as.data.frame(apply(pts2, 2, function(x) c(min(x)-5, max(x)+5)))
ini3D(argsPlot3d = list(xlim = lim2$z1, ylim = lim2$z2, zlim = lim2$z3))
plotCones3D(
  pts2[,1:3],
  drawPoint = T,
  argsPolygon3d = list(alpha = 1, color = "red"),
  rectangle = TRUE
)
plotCones3D(
  pts2[,1:3],
  argsPolygon3d = list(alpha = 1, color = "green"),
  direction = -1,
  rectangle = TRUE
)
finalize3D()
rglwidget()
```

```{r color scales, include=FALSE}
# https://cran.r-project.org/web/packages/ggsci/vignettes/ggsci.html
mypal <- pal_npg("nrc", alpha = 0.7)(9)
mypal <- pal_simpsons()(9)
# mypal
library("scales")
show_col(mypal)
```

```{r Define theme and color scales, include=FALSE, eval=FALSE}
theme_publish <- function() {
  library(grid)
  library(ggthemes)
  return(theme_foundation(base_size = 10) +
    theme(
      # plot.title = element_text(face = "bold", size = rel(1.2), hjust = 0.5),
      text = element_text(face = "plain"),
      # panel.background = element_rect(colour = NA),
      panel.spacing = unit(0.5, "cm"),
      plot.background = element_blank(),
      # panel.border = element_rect(colour = NA),
      axis.title = element_text(face = "plain",size = rel(1)),
      # axis.title.y = element_text(angle=90,vjust =2),
      # axis.title.x = element_text(vjust = -0.2),
      axis.text = element_text(),
      axis.line = element_line(colour="black"),
      axis.ticks = element_line(),
      panel.grid.major = element_line(colour="#f0f0f0"),
      panel.grid.minor = element_blank(),
      legend.key = element_rect(colour = NA),
      legend.position = "bottom",
      legend.direction = "horizontal",
      # legend.key.size = unit(0.2, "cm"),
      legend.key.height = unit(0.2, "cm"), 
      legend.key.width = unit(0.75, "cm"),
      # legend.margin = unit(0, "cm"),
      legend.title = element_blank(), #element_text(face="italic"),
      # plot.margin=unit(c(10,5,5,5),"mm"),
      # strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
      # strip.text = element_text(face="bold"),
    )
  )
}





# display.brewer.pal(n = 11, name = "RdYlBu")
pal <- brewer.pal(n = 11, name = "RdYlBu")
# algConfigs <- algConfigs[c(2,5, 3,6, 1,4, 8,11, 9,12, 7,10)]
palAlgConfigs <- wes_palette("Zissou1", algConfigsN, type = "continuous")
palAlgConfigs <- c(brewer.pal(n =  algConfigsN/2, name = "Reds"), brewer.pal(n =  algConfigsN/2, name = "Blues"))

scale_color_algConfig <- scale_color_manual(
  values = setNames(palAlgConfigs, algConfigs),
  drop = F)
scale_fill_algConfig <- scale_fill_manual(
  values = setNames(palAlgConfigs, algConfigs),
  drop = F)

scale_color_nodesel_varsel <- scale_color_manual(
  values = setNames(pal[c(1,3,9,11)], nodeselVarselConfigs),
  drop = F)

scale_color_ob <- scale_color_manual(
  values = setNames(pal[c(1,3,9,11)], c("C", "F", "N", "OSS")),
  drop = F)

scale_fill_ob <- scale_fill_manual(
  values = setNames(pal[c(1,3,9,11)], c("C", "F", "N", "Unsolved")),
  drop = F)

scale_linetype_ob <- scale_linetype_manual(
  values = c("C" = 1, "F" = 2, "N" = 3),
  drop = F)

# scale_color_nodesel <- scale_color_manual(
#   values = c("breadth" = "red", "depth" = "green"),
#   drop = F)

# scale_alpha_varsel <- scale_alpha_manual(
#   values = c("mof" = 1, "mfavg" = 0.75),
#   drop = F)

# scale_linetype_valsel <- scale_linetype_manual(
#   values = c("mof" = 1, "mfavg" = 2),
#   drop = F)
  

scale_linetype_nodesel_varsel <- scale_linetype_manual(
  values = c("B" = 1, "D" = 2, "OSS" = 3, "Unsolved" = 1),
  drop = F)
```

Given a lower bound set $L$ (the line segments) and an upper bound set $U$ with local upper bounds $N(U)$, the search region $A$ is given by the closure of the gray area:

```{r}
bBox <- list(m = c(0,0), M = c(12,14))
lb <- 
  tibble(z1 = c(1, 2, 4, 8), 
         z2 = c(8, 4, 2, 1))
ub <- 
  tibble(z1 = c(2, 4, 6, 9), 
         z2 = c(7, 4, 3, 2))
lUb <- 
  tibble(z1 = c(2,  4, 6, 9, 12), 
         z2 = c(14, 7, 4, 3, 2 ))
conesLB <- plotHull2D(lb, drawPoints = F, addRays = T, m = bBox$m, M = bBox$M, drawPlot = F, drawLines = F,
                   argsGeom_polygon = list(fill = mypal[2], alpha = 0.5)) 
ptsUB <- geom_point(aes(x = z1, y = z2, shape = "rou"), data = ub, show.legend = T) 
conesLUB <- plotCones2D(lUb, drawPlot = F, m = bBox$m, M = bBox$M, drawLines = F, direction = -1,
                     drawPoint = F, argsGeom_polygon = list(fill = mypal[1], alpha = 0.5))
lbLines <- geom_line(aes(x = z1, y = z2), data = lb)
ptsLUB <- geom_point(aes(x = z1, y = z2, shape = "sq"), data = lUb, show.legend = T)
layO <- ggplot() +
  xlim(bBox$m[1], bBox$M[1]) + ylim(bBox$m[2], bBox$M[2]) +
  gMOIPTheme() +
  theme(
    legend.position = c(0.85,0.85),
    legend.background = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()) +
  xlab("z1") + ylab("z2")

p <- layO + conesLUB + conesLB + ptsUB + ptsLUB + lbLines +
  geom_text(aes(2.75,5.5, label = "A")) + 
  geom_text(aes(lb$z1[1], lb$z2[1]-2, label = "L"), nudge_x = 0.15) + 
  scale_shape_manual(
    values = c(rou = 19, sq = 15), 
    labels = c("U", "N(U)"), name = NULL)
p  
```

A lower bound set (solid and dashed lines) partially dominated by the upper bound set. The dominated area of the lower bound set is represented by the dashed lines. Three disjoint subproblems are created in the objective space by applying objective branching on $y1$, $y2$ and $y3$, represented by the red circles. The constraints added when applying objective branching are represented by the dotted lines. A new non-dominated point feasible for the corresponding problem can only be found in one of these subproblems (gray areas):

```{r}
bBox <- list(m = c(0,0), M = c(14,11))
lb <- 
  tibble(z1 = c(2, 4, 13), 
         z2 = c(9, 5, 0.5))
lUb <- 
  tibble(z1 = c(1,  3, 5, 7, 8, 9, 14), 
         z2 = c(11, 9, 6, 5, 3, 2, 1))
ub <- 
  tibble(z1 = c(1, 3, 5, 7, 8, 9), 
         z2 = c(9, 6, 5, 3, 2, 1))
ob <-   
  tibble(z1 = c(3, 7, 14), 
         z2 = c(9, 6, 1))
ptsOB <- geom_point(aes(x = z1, y = z2), data = ob, shape = 21, size = 5, color = mypal[5])
ptsOBText <- geom_text(aes(x = z1, y = z2), data = ob, label = c("y1", "y2", "y3"), nudge_y = 0.5)
conesOB <- plotCones2D(ob, drawPlot = F, m = bBox$m, M = bBox$M, drawLines = T, direction = -1,
                     drawPoint = F, argsGeom_polygon = list(fill = NA, alpha = 0.5), argsGeom_path = list(linetype = 3, size = 0.9))
conesLB <- plotHull2D(lb, drawPoints = F, addRays = T, m = bBox$m, M = bBox$M, drawPlot = F, drawLines = F,
                   argsGeom_polygon = list(fill = mypal[2], alpha = 0.5)) 
ptsUB <- geom_point(aes(x = z1, y = z2, shape = "rou"), data = ub, show.legend = T) 
conesLUB <- plotCones2D(lUb, drawPlot = F, m = bBox$m, M = bBox$M, drawLines = F, direction = -1,
                     drawPoint = F, argsGeom_polygon = list(fill = mypal[1], alpha = 0.5))
lb1 <- tibble(z1 = c(12,13), z2 = 7 - 0.5*c(12,13))
lbLines1 <- geom_line(aes(x = z1, y = z2), data = lb1, linetype = 1)
lb2 <- tibble(z1 = c(3.5, 4), z2 = 13 - 2*c(3.5, 4))
lbLines2 <- geom_line(aes(x = z1, y = z2), data = lb2, linetype = 1)
lb3 <- tibble(z1 = c(4,7), z2 = 7 - 0.5*c(4,7))
lbLines3 <- geom_line(aes(x = z1, y = z2), data = lb3, linetype = 1)
lb4 <- tibble(z1 = c(2, 3), z2 = 13 - 2*c(2, 3))
lbLines4 <- geom_line(aes(x = z1, y = z2), data = lb4, linetype = 1)
lb5 <- tibble(z1 = c(7,12), z2 = 7 - 0.5*c(7,12))
lbLines5 <- geom_line(aes(x = z1, y = z2), data = lb5, linetype = 2)
lb6 <- tibble(z1 = c(3, 3.5), z2 = 13 - 2*c(3, 3.5))
lbLines6 <- geom_line(aes(x = z1, y = z2), data = lb6, linetype = 2)
ptsLUB <- geom_point(aes(x = z1, y = z2, shape = "sq"), data = lUb, show.legend = T)
layO <- ggplot() +
  xlim(bBox$m[1], bBox$M[1]) + ylim(bBox$m[2], bBox$M[2]) +
  gMOIPTheme() +
  theme(
    legend.position = c(0.85,0.85),
    legend.background = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()) +
  xlab("z1") + ylab("z2")
p <- layO + conesLUB + conesLB + conesOB + ptsUB + ptsOB + ptsLUB + ptsOBText +
  lbLines1 + lbLines2 + lbLines3 + lbLines4 + lbLines5 + lbLines6 +
  scale_shape_manual(
    values = c(rou = 19, sq = 15), 
    labels = c("U", "N(U)"), name = NULL)
p  
```

Bi-objective case: The dominated part of the lower bound set is represented by the dashed line. When objective branching is applied on each local upper bound, the search areas (gray) are disjoint:

```{r}
bBox <- list(m = c(0,0), M = c(7,7))
lb <- 
  tibble(z1 = c(1, 6), 
         z2 = c(6, 1))
lUb <- 
  tibble(z1 = c(3, 7), 
         z2 = c(7, 3))
ub <- 
  tibble(z1 = c(3), 
         z2 = c(3))
ob <-   
  tibble(z1 = c(3, 3, 7), 
         z2 = c(7, 3, 3))
lb1 <- tibble(z1 = c(1,3), z2 = 7 - c(1,3))
lb2 <- tibble(z1 = c(3,4), z2 = 7 - c(3,4))
lb3 <- tibble(z1 = c(4,6), z2 = 7 - c(4,6))
conesOB <- plotCones2D(ob, drawPlot = F, m = bBox$m, M = bBox$M, drawLines = T, direction = -1,
                     drawPoint = F, argsGeom_polygon = list(fill = NA, alpha = 0.5), argsGeom_path = list(linetype = 3, size = 0.5))
conesLB <- plotHull2D(lb, drawPoints = F, addRays = T, m = bBox$m, M = bBox$M, drawPlot = F, drawLines = F,
                   argsGeom_polygon = list(fill = mypal[2], alpha = 0.5)) 
ptsUB <- geom_point(aes(x = z1, y = z2, shape = "rou"), data = ub, show.legend = T) 
conesLUB <- plotCones2D(lUb, drawPlot = F, m = bBox$m, M = bBox$M, drawLines = F, direction = -1,
                     drawPoint = F, argsGeom_polygon = list(fill = mypal[1], alpha = 0.5))
lbLines1 <- geom_line(aes(x = z1, y = z2), data = lb1, linetype = 1)
lbLines2 <- geom_line(aes(x = z1, y = z2), data = lb2, linetype = 2)
lbLines3 <- geom_line(aes(x = z1, y = z2), data = lb3, linetype = 1)
ptsLUB <- geom_point(aes(x = z1, y = z2, shape = "sq"), color = mypal[8], data = lUb)
layO <- ggplot() +
  xlim(bBox$m[1], bBox$M[1]) + ylim(bBox$m[2], bBox$M[2]) +
  gMOIPTheme() +
  theme(
    legend.background = element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()) +
  xlab("z1") + ylab("z2")
p <- layO + conesLUB + conesLB + conesOB + ptsUB + ptsLUB + 
  lbLines1 + lbLines2 + lbLines3 +
  scale_shape_manual(
    values = c(rou = 19, sq = 19), 
    labels = c("U", "N(U)"), name = NULL)
p  
```

Three-objective case: The dominated part of the lower bound set is the blue area in the middle. When objective branching is applied on each local upper bound, there are redundancies between the subproblems (each subproblem defines a gray search area). Every point in the red areas is included in the search area of more than one subproblem:

```{r}
ubSet <- matrix( c(1,1,1), ncol = 3, byrow = TRUE)
row.names(ubSet) <- paste0("z", 1)
lim <- c(0,7)  # bounding box for problem
m <- rep(lim[1],3)
M <-rep(lim[2],3)
augment <- function(idx, a, b, prefix = "") {
  res <- NULL
  for (i in idx) {
    c <- b
    c[i] <- a[i]
    res <- rbind(res,c)
  }
  row.names(res) <- paste0(prefix,idx)
  return(res)
}
zHat <- augment(1:3, M, m)
# iteration 0
U <- matrix(M, ncol = 3, byrow = T)
row.names(U) <- "u0"
N <- NULL
NHat <- rbind(N,zHat)
z <- ubSet[1,1:3]
u0 <- M
U <- augment(1:3, z, u0, "u")
N<- rbind(N, ubSet[1,1:3, drop=F])

view <- matrix( c(-0.640044808387756, -0.205455526709557, 0.740358412265778, 0, 0.767741799354553, -0.208960145711899, 0.605729579925537, 0, 0.0302549730986357, 0.956098079681396, 0.29148057103157, 0, 0, 0, 0, 1), nc = 4)

lim <- c(lim[1]-0.05,lim[2]+0.05)
## all relevant points for drawing search areas
vtx <- matrix(c(
  1,1,4,
  1,4,1,
  4,1,1,
  0,0,7,
  0,7,0,
  7,0,0,
  0,0,6,
  0,6,0,
  6,0,0,
  0,1,5,
  0,5,1,
  1,0,5,
  1,5,0,
  5,0,1,
  5,1,0,
  1,1,7,
  1,7,1,
  7,1,1,
  7,1,0,
  7,0,1,
  1,7,0,
  1,0,7,
  0,7,1,
  0,1,7,
  7,7,0,
  7,0,7,
  0,7,7,
  7,1,7,
  7,7,1,
  1,7,7
), ncol = 3, byrow = T)
vtx <- vtx %>% set_colnames(c("x", "y", "z")) %>% as_tibble() %>% arrange(x,y,z)
ini3D(argsPlot3d = list(xlim = lim, ylim = lim, zlim = lim))
# Points
plotPoints3D(N, argsPlot3d = list(type="p", col = "black", size = 10))
plotPoints3D(U, argsPlot3d = list(type="p", col = mypal[8], size = 10), argsText3d = list(col = "blue", adj = c(1,1)))
# Plane
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPlanes3d = list(alpha = 0.75, col = mypal[2]))
plotPlane3D(c(1,1,1), point = c(2,2,2), argsLines = list(col = "black", lines = 100, alpha = 0.75), useLines = T, useShade = F)
# Overlapping areas
args <- list(color = mypal[5], alpha = 0.9)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,x,z)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(z,y,x)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
# Non-overlapping areas
args <- list(color = mypal[3], alpha = 0.4)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,x,z)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,z,x)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
finalize3D(argsAxes3d = list(edges = c("x", "y", "z"), box = F, labels=FALSE, tick=FALSE))
# finalize3D(argsAxes3d = list(edges = "bbox"))
view <- matrix( c(-0.601261854171753, -0.375613540410995, 0.705264508724213, 0, 0.79587709903717, -0.36011865735054, 0.48671817779541, 0, 0.071161076426506, 0.853949129581451, 0.515467822551727, 0, 0, 0, 0, 1), nc = 4)
loadView(v = view, clear = F, zoom = 0.8)
rglwidget()
```


Lower bound set (hyperplane) partially dominated by the upper bound set (black points with red dominance cones). Note the two disjoint search areas above the hyperplane:

```{r}
pts <- tibble::tribble(
   ~z1, ~z2, ~z3,  ~nd,
    -2,   2,   5, TRUE,
    -2,   4,   4, TRUE,
    -2,   5,   3, TRUE,
    -1,   1,   6, TRUE,
    -1,   6,   2, TRUE,
     0,   0,   6, TRUE,
     0,   6,   0, TRUE,
     1,  -1,   6, TRUE,
     1,   6,  -1, TRUE,
     2,   5,  -2, TRUE,
     3,  -2,   5, TRUE,
     4,  -2,   4, TRUE,
     4,   4,  -2, TRUE,
     5,  -2,   3, TRUE,
     5,   3,  -2, TRUE,
     6,  -1,   1, TRUE,
     6,   0,   0, TRUE,
     6,   2,  -1, TRUE
   )

## First plot
ini3D(argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10), 
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10), 
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
lub <- Rfast::colMinsMaxs(as.matrix(pts[,1:3]))[2,]
M <- rgl::par3d()$bbox
lub <- rbind(lub, c(M[2],M[4],M[6])-0.01)
llb <- c(M[1],M[3],M[5])+ 0.01
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[8]), drawPoint = T, rectangle = T, argsPlot3d = list(size = 10))
plotCones3D(lub[2,], direction = -1, argsPolygon3d = list(alpha = 0.4, color = mypal[1]), drawPoint = F)
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.6, col = mypal[7]))
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(llb, argsPolygon3d = list(alpha = 0.75, color = mypal[3]), drawPoint = F)
clipplanes3d(1, 1, 1, -9)
useSubscene3d(root)
finalize3D(argsAxes3d = list(edges = "bbox"), argsTitle3d = list(xlab = "", ylab = "", zlab = ""))
view <- matrix( c(0.710081100463867, -0.417695313692093, -0.566847920417786, 0, -0.695306360721588, -0.288975149393082, -0.658059477806091, 0, 0.111063443124294, 0.861408114433289, -0.495621830224991, 0, 0, 0, 0, 1), nc = 4)
loadView(v = view, clear = FALSE, zoom = 0.7)
rglwidget()
```

Objective branching applied to $y^1$ and $y^2$. One subproblem is fully included in the other subproblem:

```{r}
## Second plot
ini3D(clear = T, argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10), 
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10), 
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))

plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.6, col = mypal[7]))
plotCones3D(lub, direction = -1, argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T)
plotmath3d(lub-1, text = expression(y^1, y^2), cex = 1.2, font = 3)
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[8]), drawPoint = T, drawLines = T, argsPlot3d = list(size = 10))
clipplanes3d(-1, -1, -1, 9)
useSubscene3d(root)
finalize3D(argsAxes3d = list(edges = "bbox"), argsTitle3d = list(xlab = "", ylab = "", zlab = ""))
view <- matrix( c(0.860876560211182, -0.30657571554184, -0.406083613634109, 0, -0.448700070381165, -0.081096462905407, -0.889995634555817, 0, 0.239919260144234, 0.948385238647461, -0.207374647259712, 0, 0, 0, 0, 1), nc = 4)
loadView(v = view, clear = FALSE, zoom = 0.7)
rglwidget()
```

Another view:

```{r}
## First plot
id <- open3d()
mfrow3d(nr = 1, nc = 2, sharedMouse = F)
ini3D(argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10),
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10),
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
lub <- Rfast::colMinsMaxs(as.matrix(pts[,1:3]))[2,]
M <- rgl::par3d()$bbox
lub <- rbind(lub, c(M[2],M[4],M[6])-0.01)
llb <- c(M[1],M[3],M[5])+ 0.01
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T, rectangle = T)
plotCones3D(lub[2,], direction = -1, argsPolygon3d = list(alpha = 0.4, color = mypal[7]), drawPoint = F)
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.4, col = mypal[7]))
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(llb, argsPolygon3d = list(alpha = 0.4, color = mypal[4]), drawPoint = F)
# plotHull3D(unique(pts))
clipplanes3d(1, 1, 1, -9)
useSubscene3d(root)
axes3d()
next3d()
## Second plot
ini3D(clear = FALSE, argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10),
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10),
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
# plotPoints3D(pts)
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.8, col = mypal[7]))
plotCones3D(lub[1,], direction = -1, argsPolygon3d = list(alpha = 1, color = mypal[4]), drawPoint = T)
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T, drawLines = F)
useSubscene3d(root)
axes3d()
rglwidget()
close3d()
```

```{r, include=F}
rm(list = ls(all.names = TRUE))
```