---
title: "3D plots in tri-objective programming"
author: "Lars Relund Nielsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3D plots in tri-objective programming}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(knitr)
library(rgl)
library(ggsci)
library(tidyverse)
library(magrittr)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
rgl::par3d("family" = "serif")
opts_chunk$set(
  collapse = TRUE,
  
  comment = "#>",
  warning=FALSE, message=FALSE, include = TRUE, 
  cache = TRUE,
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)
```

The `gMOIP` package can be used to make 2D and 3D plots of linear programming (LP), integer linear
programming (ILP), or mixed integer linear programming (MILP) models with up to three objectives.
This include the polytope, integer points, ranges and iso profit curve. Plots of both the solution
and criterion space are possible. For instance the non-dominated (Pareto) set for bi-objective
LP/ILP/MILP programming models.

The package also include an `inHull` function for checking if a set of points is inside/at/outside
the convex hull of a set of vertices (for arbitrary dimension).

Finally, the package also contains functions for generating (non-dominated) points in $\mathbb{R}_n$
and classifying non-dominated points as supported extreme, supported non-extreme and unsupported.


## Installation

Install the latest stable release from CRAN:

``` r
install.packages("gMOIP")
```

Alternatively, install the latest development version from GitHub:

``` r
install.packages("devtools")
devtools::install_github("relund/gMOIP")
library(gMOIP)
```

```{r, cache=FALSE, include=FALSE}
library(gMOIP)
```

## Drawing shapes related to MO

```{r color scales, include=FALSE}
mypal <- pal_npg("nrc", alpha = 0.7)(9)
mypal <- pal_simpsons()(9)
# mypal
library("scales")
# show_col(mypal)
```

Complex shapes can be drawn using cones, hulls etc. Plot of a single non-dominated point and its corresponding local upper bounds given the bounding box. Green areas are overlapping if branch on each local upper bound.

```{r Fig3 in Forget20, webgl = TRUE, echo=FALSE, cache=FALSE}
ubSet <- matrix( c(1,1,1), ncol = 3, byrow = TRUE)
row.names(ubSet) <- paste0("z", 1)
lim <- c(0,7)  # bounding box for problem
m <- rep(lim[1],3)
M <-rep(lim[2],3)
augment <- function(idx, a, b, prefix = "") {
  res <- NULL
  for (i in idx) {
    c <- b
    c[i] <- a[i]
    res <- rbind(res,c)
  }
  row.names(res) <- paste0(prefix,idx)
  return(res)
}
zHat <- augment(1:3, M, m)
# iteration 0
U <- matrix(M, ncol = 3, byrow = T)
row.names(U) <- "u0"
N <- NULL
NHat <- rbind(N,zHat)
z <- ubSet[1,1:3]
u0 <- M
U <- augment(1:3, z, u0, "u")
N<- rbind(N, ubSet[1,1:3, drop=F])

view <- matrix( c(-0.640044808387756, -0.205455526709557, 0.740358412265778, 0, 0.767741799354553, -0.208960145711899, 0.605729579925537, 0, 0.0302549730986357, 0.956098079681396, 0.29148057103157, 0, 0, 0, 0, 1), nc = 4)

rgl::mfrow3d(nr = 1, nc = 2, sharedMouse = F)

lim <- c(lim[1]-0.05,lim[2]+0.05)
## First plot
ini3D(argsPlot3d = list(xlim = lim, ylim = lim, zlim = lim))
# plotRectangle3D(m,M)
plotPoints3D(N, addText = "coord", argsPlot3d = list(type="p", col = "black", size = 8))
plotPoints3D(U, addText = "coord", argsPlot3d = list(type="p", col = mypal[3], size = 8), argsText3d = list(col = "blue", adj = c(1,1)))
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPersp3d = list(col = mypal[2]), usePersp3d = T, lines = 50)
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPlanes3d = list(alpha = 0.7, col = mypal[8]))
plotCones3D(U[1:3,], drawPoint = F, argsPolygon3d = list(alpha =0.3, color = mypal[1]), rectangle = F, direction = -1, drawLines = F)
finalize3D(argsAxes3d = list(edges = "bbox", box = F))
#rglwidget()


loadView(v = view, clear = FALSE)

## all relevant points for drawing search areas
vtx <- matrix(c(
  1,1,4,
  1,4,1,
  4,1,1,
  0,0,7,
  0,7,0,
  7,0,0,
  0,0,6,
  0,6,0,
  6,0,0,
  0,1,5,
  0,5,1,
  1,0,5,
  1,5,0,
  5,0,1,
  5,1,0,
  1,1,7,
  1,7,1,
  7,1,1,
  7,1,0,
  7,0,1,
  1,7,0,
  1,0,7,
  0,7,1,
  0,1,7,
  7,7,0,
  7,0,7,
  0,7,7,
  7,1,7,
  7,7,1,
  1,7,7
), ncol = 3, byrow = T)
vtx <- vtx %>% set_colnames(c("x", "y", "z")) %>% as_tibble() %>% arrange(x,y,z)
# plotPoints3D(vtx, addText = "coord")
# print(vtx, n=100)

## Second plot
ini3D(clear = FALSE, argsPlot3d = list(xlim = lim, ylim = lim, zlim = lim))
plotPoints3D(N, addText = "coord", argsPlot3d = list(type="p", col = "black", size = 8))
plotPoints3D(U, addText = "coord", argsPlot3d = list(type="p", col = mypal[7], size = 8), argsText3d = list(col = "blue", adj = c(1,1)))
# plotPlane3D(c(1,1,1), point = c(2,2,2), argsPersp3d = list(col = mypal[2]), usePersp3d = T, lines = 50)
plotPlane3D(c(1,1,1), point = c(2,2,2), argsPlanes3d = list(alpha = 0.3, col = mypal[8]))

args <- list(color = mypal[9], alpha = 0.7)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,x,z)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(z,y,x)
plotHull3D(vtx[c(19:24,26:27),], drawPolygons = T, argsPolygon3d = args)

args <- list(color = mypal[4], alpha = 0.2)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,x,z)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
vtx <- vtx %>% arrange(y,z,x)
plotHull3D(vtx[c(10,11,12,13,19,20,24,25,27,28),] , drawPolygons = T, argsPolygon3d = args)
finalize3D(argsAxes3d = list(edges = "bbox", box = F))
loadView(v = view, clear = FALSE)
```

A set of non-dominated points and the areas they dominate. Light blue plane are a lower bound.

```{r Fig4, webgl = TRUE, echo=FALSE, cache=FALSE}
pts <- data.frame(
          z1 = c(-2, -2, -2, -1, -1, 0, 0, 1, 1, 2, 3, 4, 4, 5, 5, 6, 6, 6),
          z2 = c(2, 4, 5, 1, 6, 0, 6, -1, 6, 5, -2, -2, 4, -2, 3, -1, 0, 2),
          z3 = c(5, 4, 3, 6, 2, 6, 0, 6, -1, -2, 5, 4, -2, 3, -2, 1, 0, -1),
          nd = c(TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,
                 TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE,TRUE)
)

## First plot
open3d()
mfrow3d(nr = 1, nc = 2, sharedMouse = F)
ini3D(argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10),
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10),
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
lub <- Rfast::colMinsMaxs(as.matrix(pts[,1:3]))[2,]
M <- rgl::par3d()$bbox
lub <- rbind(lub, c(M[2],M[4],M[6])-0.01)
llb <- c(M[1],M[3],M[5])+ 0.01
# plotPoints3D(pts)
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T, rectangle = T)
plotCones3D(lub[2,], direction = -1, argsPolygon3d = list(alpha = 0.4, color = mypal[7]), drawPoint = F)
# plotHull3D(unique(pts))
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.4, col = mypal[7]))
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(llb, argsPolygon3d = list(alpha = 0.4, color = mypal[4]), drawPoint = F)
# plotHull3D(unique(pts))
clipplanes3d(1, 1, 1, -9)
useSubscene3d(root)
axes3d()

## Second plot
ini3D(clear = FALSE, argsPlot3d = list(xlim = c(min(pts$z1)-2,max(pts$z1)+10),
                        ylim = c(min(pts$z2)-2,max(pts$z2)+10),
                        zlim = c(min(pts$z3)-2,max(pts$z3)+10)))
# plotPoints3D(pts)
plotPlane3D(c(1,1,1), point = c(3,3,3), argsPlanes3d = list(alpha = 0.8, col = mypal[7]))
plotCones3D(lub[1,], direction = -1, argsPolygon3d = list(alpha = 1, color = mypal[4]), drawPoint = T)
root <- currentSubscene3d()
newSubscene3d(
  "inherit",
  "inherit",
  "inherit",
  copyShapes = T,
  copyBBoxDeco = F,
  parent = root
)
plotCones3D(pts[pts$nd,c("z1","z2","z3")], argsPolygon3d = list(alpha = 1, color = mypal[1]), drawPoint = T, drawLines = F)
useSubscene3d(root)
axes3d()
```

```{r, include=F}
rgl.quit() 
rm(list = ls(all.names = TRUE))
```