---
title: "Plotting polytopes in 3D - Example 3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting polytopes in 3D - Example 3}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(knitr)
library(rgl)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
opts_chunk$set(
  collapse = TRUE,
  #comment = "#>",
  warning=FALSE, message=FALSE, include = TRUE, 
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)
```

With `gMOIP` you can make 3D plots of the polytope/feasible region/solution space of a linear programming (LP), integer linear programming (ILP) model, or mixed integer linear programming (MILP) model. This vignette gives examples on how to make plots given a model with three variables. 

First we load the package:

```{r setup}
library(gMOIP)
```

We define the model:

```{r ex3Model}
A <- matrix( c(
   -1, 1, 0,
   1, 4, 0,
   2, 1, 0,
   3, -4, 0,
   0, 0, 4
), nc = 3, byrow = TRUE)
b <- c(5, 45, 27, 24, 10)
obj <- c(5, 45, 15)
```

We load the preferred view angle for the RGL window:

```{r ex3View}
view <- matrix( c(0.976349174976349, -0.202332556247711, 0.0761845782399178, 0, 0.0903248339891434,
                  0.701892614364624, 0.706531345844269, 0, -0.196427255868912, -0.682940244674683,
                  0.703568696975708, 0, 0, 0, 0, 1), nc = 4)
```

The LP polytope:

```{r ex3LP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj, labels = "coord")
```

Note you can zoom/turn/twist the figure with your mouse (`rglwidget`).

The ILP model (note since the vertices are integer the LP and ILP faces are equal):

```{r ex3ILP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","i"), plotOptimum = TRUE, obj = obj, 
             argsTitle3d = list(main = "With LP faces"), argsPlot3d = list(box = F, axes = T) )
```


Let us have a look at some MILP models (we use static images). MILP model with variable 1 and 3 
integer:

```{r ex3MILP_1, webgl = TRUE}
loadView(v = view, close = T, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","c","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 2 and 3 integer:

```{r ex3MILP_2, webgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("c","i","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 and 2 integer:

```{r ex3MILP_3, webgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","c"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 integer:

```{r ex3MILP_4, webgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 2 integer:

```{r ex3MILP_5, webgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 3 integer:

```{r ex3MILP_6, webgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","c","i"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

```{r, include=F, eval=FALSE}
detach(package:gMOIP, unload=TRUE)
detach(package:rgl, unload=TRUE)
```