---
title: "Plotting polytopes in 3D - Example 4"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting polytopes in 3D - Example 4}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(knitr)
library(rgl)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
opts_chunk$set(
  collapse = TRUE,
  webgl = TRUE,
  #comment = "#>",
  warning=FALSE, message=FALSE, include = TRUE, 
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)
```

With `gMOIP` you can make 3D plots of the polytope/feasible region/solution space of a linear programming (LP), integer linear programming (ILP) model, or mixed integer linear programming (MILP) model. This vignette gives examples on how to make plots given a model with three variables. 

First we load the package:

```{r setup}
library(gMOIP)
```

We define the model:

```{r ex4Model}
Ab <- matrix( c(
   1, 1, 2, 5,
   2, -1, 0, 3,
   -1, 2, 1, 3,
   0, -3, 5, 2
   #   0, 1, 0, 4,
   #   1, 0, 0, 4
), nc = 4, byrow = TRUE)
A <- Ab[,1:3]
b <- Ab[,4]
obj <- c(1,1,3)
```

We load the preferred view angle for the RGL window:

```{r ex4View}
view <- matrix( c(-0.452365815639496, -0.446501553058624, 0.77201122045517, 0, 0.886364221572876,
                  -0.320795893669128, 0.333835482597351, 0, 0.0986008867621422, 0.835299551486969,
                  0.540881276130676, 0, 0, 0, 0, 1), nc = 4)
```

The LP polytope:

```{r ex4LP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj, labels = "coord")
```

Note in the html file you can zoom/turn/twist the figure with your mouse (`rglwidget`).

The ILP model with LP and ILP faces:

```{r ex4ILP, webgl = TRUE}
loadView(v = view)
rgl::mfrow3d(nr = 1, nc = 2, sharedMouse = TRUE)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","i"), plotOptimum = TRUE, obj = obj, 
             argsTitle3d = list(main = "With LP faces"), argsPlot3d = list(box = F, axes = T) )
plotPolytope(A, b, faces = c("i","i","i"), type = c("i","i","i"), plotFeasible = FALSE, obj = obj,
             argsTitle3d = list(main = "ILP faces") )
```

Let us have a look at some MILP models. MILP model with variable 1 and 3 
integer:

```{r ex4MILP_1}
loadView(v = view, close = T, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","i"), plotOptimum = TRUE, obj = obj, faces = c("c","c","c"))
```

MILP model with variable 2 and 3 integer:

```{r ex4MILP_2}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","i"), plotOptimum = TRUE, obj = obj, faces = c("c","c","c"))
```

MILP model with variable 1 and 2 integer:

```{r ex4MILP_3}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","i","c"), plotOptimum = TRUE, obj = obj, faces = c("c","c","c"))
```

MILP model with variable 1 integer:

```{r ex4MILP_4}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","c"), plotOptimum = TRUE, obj = obj, faces = c("c","c","c"))
```

MILP model with variable 2 integer:

```{r ex4MILP_5}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","c"), plotOptimum = TRUE, obj = obj, faces = c("c","c","c"))
```

MILP model with variable 3 integer:

```{r ex4MILP_6}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","c","i"), plotOptimum = TRUE, obj = obj, faces = c("c","c","c"))
```

```{r, include=F}
rm(list = ls(all.names = TRUE))
``


