---
title: "Plotting polytopes in 3D"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting polytopes in 3D}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(knitr)
library(rgl)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "poly3d-",
  warning=FALSE, message=FALSE, include = TRUE, 
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.62
)

knit_hooks$set(webgl = hook_webgl, rgl = hook_rgl)
```

With `gMOIP` you can make 3D plots of the polytope/feasible region/solution space of a linear programming (LP), integer linear programming (ILP) model, or mixed integer linear programming (MILP) model. This vignette gives examples on how to make plots given a model with three variables. 

First we load the package:

```{r setup}
library(gMOIP)
```

We present four examples. 

## Example 1

We define the model  $\max \{cx | Ax \leq b\}$ (could also be minimized) with three variables:

```{r ex1Model}
A <- matrix( c(
   3, 2, 5,
   2, 1, 1,
   1, 1, 3,
   5, 2, 4
), nc = 3, byrow = TRUE)
b <- c(55, 26, 30, 57)
obj <- c(20, 10, 15)
```

We load the preferred view angle for the RGL window:

```{r ex1View}
view <- matrix( c(-0.412063330411911, -0.228006735444069, 0.882166087627411, 0, 0.910147845745087,
                  -0.0574885793030262, 0.410274744033813, 0, -0.042830865830183, 0.97196090221405,
                  0.231208890676498, 0, 0, 0, 0, 1), nc = 4)
```

The LP polytope:

```{r ex1LP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj)
```

Note in the html file you can zoom/turn/twist the figure with your mouse (`rglwidget`).

The ILP model with LP and ILP faces:

```{r ex1ILP, webgl = TRUE}
loadView(v = view)
mfrow3d(nr = 1, nc = 2, sharedMouse = TRUE)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","i"), plotOptimum = TRUE, obj = obj, 
             argsTitle3d = list(main = "With LP faces"), argsPlot3d = list(box = F, axes = T) )
plotPolytope(A, b, faces = c("i","i","i"), type = c("i","i","i"), plotFeasible = FALSE, obj = obj,
             argsTitle3d = list(main = "ILP faces") )
```

Let us have a look at some MILP models (we use static images). MILP model with variable 1 and 3 
integer:

```{r ex1MILP_1, rgl = TRUE}
loadView(v = view, close = T, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","c","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 2 and 3 integer:

```{r ex1MILP_2, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("c","i","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 and 2 integer:

```{r ex1MILP_3, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","c"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 integer:

```{r ex1MILP_4, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 2 integer:

```{r ex1MILP_5, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 3 integer:

```{r ex1MILP_6, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","c","i"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```


## Example 2

We define the model:

```{r ex2Model}
A <- matrix( c(
   1, 1, 1,
   3, 0, 1
), nc = 3, byrow = TRUE)
b <- c(10, 24)
obj <- c(20, 10, 15)
```

We load the preferred view angle for the RGL window:

```{r ex2View}
view <- matrix( c(-0.812462985515594, -0.029454167932272, 0.582268416881561, 0, 0.579295456409454,
                  -0.153386667370796, 0.800555109977722, 0, 0.0657325685024261, 0.987727105617523,
                  0.14168381690979, 0, 0, 0, 0, 1), nc = 4)
```

The LP polytope:

```{r ex2LP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj, labels = "coord")
```

Note in the html file you can zoom/turn/twist the figure with your mouse (`rglwidget`).

The ILP model with LP and ILP faces:

```{r ex2ILP, webgl = TRUE}
loadView(v = view)
mfrow3d(nr = 1, nc = 2, sharedMouse = TRUE)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","i"), plotOptimum = TRUE, obj = obj, 
             argsTitle3d = list(main = "With LP faces"), argsPlot3d = list(box = F, axes = T) )
plotPolytope(A, b, faces = c("i","i","i"), type = c("i","i","i"), plotFeasible = FALSE, obj = obj,
             argsTitle3d = list(main = "ILP faces") )
```

Let us have a look at some MILP models (we use static images). MILP model with variable 1 and 3 
integer:

```{r ex2MILP_1, rgl = TRUE}
loadView(v = view, close = T, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","c","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 2 and 3 integer:

```{r ex2MILP_2, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("c","i","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 and 2 integer:

```{r ex2MILP_3, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","c"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 integer:

```{r ex2MILP_4, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 2 integer:

```{r ex2MILP_5, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 3 integer:

```{r ex2MILP_6, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","c","i"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```


## Example 3

We define the model:

```{r ex3Model}
A <- matrix( c(
   -1, 1, 0,
   1, 4, 0,
   2, 1, 0,
   3, -4, 0,
   0, 0, 4
), nc = 3, byrow = TRUE)
b <- c(5, 45, 27, 24, 10)
obj <- c(5, 45, 15)
```

We load the preferred view angle for the RGL window:

```{r ex3View}
view <- matrix( c(0.976349174976349, -0.202332556247711, 0.0761845782399178, 0, 0.0903248339891434,
                  0.701892614364624, 0.706531345844269, 0, -0.196427255868912, -0.682940244674683,
                  0.703568696975708, 0, 0, 0, 0, 1), nc = 4)
```

The LP polytope:

```{r ex3LP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj, labels = "coord")
```

Note in the html file you can zoom/turn/twist the figure with your mouse (`rglwidget`).

The ILP model with LP and ILP faces:

```{r ex3ILP, webgl = TRUE}
loadView(v = view)
mfrow3d(nr = 1, nc = 2, sharedMouse = TRUE)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","i"), plotOptimum = TRUE, obj = obj, 
             argsTitle3d = list(main = "With LP faces"), argsPlot3d = list(box = F, axes = T) )
plotPolytope(A, b, faces = c("i","i","i"), type = c("i","i","i"), plotFeasible = FALSE, obj = obj,
             argsTitle3d = list(main = "ILP faces") )
```

Let us have a look at some MILP models (we use static images). MILP model with variable 1 and 3 
integer:

```{r ex3MILP_1, rgl = TRUE}
loadView(v = view, close = T, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","c","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 2 and 3 integer:

```{r ex3MILP_2, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("c","i","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 and 2 integer:

```{r ex3MILP_3, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","c"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 integer:

```{r ex3MILP_4, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 2 integer:

```{r ex3MILP_5, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 3 integer:

```{r ex3MILP_6, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","c","i"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```


## Example 4

We define the model:

```{r ex4Model}
Ab <- matrix( c(
   1, 1, 2, 5,
   2, -1, 0, 3,
   -1, 2, 1, 3,
   0, -3, 5, 2
   #   0, 1, 0, 4,
   #   1, 0, 0, 4
), nc = 4, byrow = TRUE)
A <- Ab[,1:3]
b <- Ab[,4]
obj = c(1,1,3)
```

We load the preferred view angle for the RGL window:

```{r ex4View}
view <- matrix( c(-0.452365815639496, -0.446501553058624, 0.77201122045517, 0, 0.886364221572876,
                  -0.320795893669128, 0.333835482597351, 0, 0.0986008867621422, 0.835299551486969,
                  0.540881276130676, 0, 0, 0, 0, 1), nc = 4)
```

The LP polytope:

```{r ex4LP, webgl = TRUE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj, labels = "coord")
```

Note in the html file you can zoom/turn/twist the figure with your mouse (`rglwidget`).

The ILP model with LP and ILP faces:

```{r ex4ILP, webgl = TRUE}
loadView(v = view)
mfrow3d(nr = 1, nc = 2, sharedMouse = TRUE)
plotPolytope(A, b, faces = c("c","c","c"), type = c("i","i","i"), plotOptimum = TRUE, obj = obj, 
             argsTitle3d = list(main = "With LP faces"), argsPlot3d = list(box = F, axes = T) )
plotPolytope(A, b, faces = c("i","i","i"), type = c("i","i","i"), plotFeasible = FALSE, obj = obj,
             argsTitle3d = list(main = "ILP faces") )
```

Let us have a look at some MILP models (we use static images). MILP model with variable 1 and 3 
integer:

```{r ex4MILP_1, rgl = TRUE}
loadView(v = view, close = T, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 2 and 3 integer:

```{r ex4MILP_2, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","i"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 and 2 integer:

```{r ex4MILP_3, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","i","c"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 1 integer:

```{r ex4MILP_4, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("i","c","c"), plotOptimum = TRUE, obj = obj, plotFaces = FALSE)
```

MILP model with variable 2 integer:

```{r ex4MILP_5, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","i","c"), plotOptimum = TRUE, obj = obj)
```

MILP model with variable 3 integer:

```{r ex4MILP_6, rgl = TRUE}
loadView(v = view, zoom = 0.75)
plotPolytope(A, b, type = c("c","c","i"), plotOptimum = TRUE, obj = obj)
```





## Making plots in LaTeX

It is possible to include the 3D plots in LaTeX using asymptote. First, save the plot as a LaTeX 
file:

```{r LaTeX, eval=FALSE}
loadView(v = view, close = F, zoom = 0.75)
plotPolytope(A, b, plotOptimum = TRUE, obj = obj, labels = "n", 
             argsTitle3d=list(xlab = '$x_1$', ylab='$y_1$', zlab='$y_2$'))
writeASY(outtype = "pdflatex", prc = TRUE, title = "polytope")  # prc = T(interactive) / F(static)
```

Next, you include the file in you LaTeX document (`test.tex`):

```{latex}
\documentclass{article}
\usepackage{a4wide}
\usepackage[pdftex]{graphicx}
\usepackage{asymptote}
\begin{document}
Example of how to include a 3D plot in a \LaTeX document:
\input{polytope}
\end{document}
```

Now, you have to run a sequence of commands:

```{bash, eval=FALSE}
pdflatex.exe --shell-escape -synctex=1 -interaction=nonstopmode test.tex
asy.exe -globalwrite test-1.asy
pdflatex.exe --shell-escape -synctex=1 -interaction=nonstopmode test.tex
```

For some plots I have experienced an error when run `asy` and the `light` function. In this case I have just removed the lines defining the `light` function in `polytope.tex`.

